---
title: "Prospective validation"
output: html_document
---

As with the previous vignettes, we begin by loading the relevant libraries and data.

```{r message=FALSE}
library( tidyverse )
library( ABCmonster )
data( MACCSbinary )
```

## Making predictions on test data

We train a model and apply it to make predictions on new data.

```{r eval=FALSE, warning=FALSE, message=FALSE}
m <- ABCtrain( MACCSbinary )      # Will use rows of MACCSbinary where Label is not-NA
P <- ABCpredict( m, MACCSbinary ) # Will use rows of MACCSbinary where Label is NA
```

The predictions are appended to the test data in `ABCpred` column. We can examine them directly by selecting the appropriate columns.

```{r eval=FALSE}
P %>% select( ABCpred, Drug, pubchem_id )
```

```{r eval=FALSE, echo=FALSE}
library( kableExtra )
P %>% select( ABCpred, Drug, pubchem_id ) %>% kable("html") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width=FALSE) %>%
    scroll_box( height="300px", box_css = "border: none; width: 35%; margin-left: auto; margin-right: auto; margin-bottom: 10px;" )
```

## Validation data

Data can be loaded directly from this package.

```{r}
data(ABCvaldata)
head(ABCvaldata)
```

```{r fig.align="center", fig.width=7, fig.height=5, out.width="100%"}
## Define labels for the "Dose" axis
doseLbl <- c("0","Top/4","Top/2","Top")

## Prepare the data for plotting
V <- ABCvaldata %>% gather( Strain, Growth, ABC16, Parental ) %>%
     mutate( Dose = match(Dose, doseLbl), DrugRep = str_c(Abbrev, Replicate) )

## Plot dose response curves
ggplot( V, aes(x=Dose, y=Growth, fill=Strain) ) +
    geom_area( alpha = 0.3, color="black", position="identity" ) +
    scale_fill_manual( values = c("ABC16"="green", "Parental"="gray50") ) +
    scale_x_continuous( breaks = 1:4, labels = doseLbl ) +
    scale_y_continuous( breaks = c(0,0.5,1) ) +
    facet_wrap( ~DrugRep, nrow=6, ncol=8 ) + theme_bw() +
    theme( axis.text.x = element_text(angle=90, vjust=0.5, hjust=1),
          strip.text.x = element_text(margin = margin(0.1,0,0.1,0,"cm")) )
```

### Concordance between replicates

```{r fig.align="center"}
## Define the color palette
pal <- c( "Parental"="tomato", "Equal"="black", "ABC16"="steelblue" )

S <- ABCvaldata %>% group_by( Drug, Abbrev, Replicate ) %>%
    summarize_at( vars(ABC16, Parental), sum ) %>%
    mutate( Value = -log2(ABC16 / Parental) ) %>% ungroup %>%
    select( -ABC16, -Parental ) %>% spread( Replicate, Value ) %>%
    mutate( Sensitivity = cut(`1`, c(-Inf, -0.1,0.1, Inf), labels=names(pal)) )
ggplot( S, aes(x=`1`,y=`2`, color=Sensitivity) ) +
    geom_point() + theme_bw() + coord_fixed() +
    scale_color_manual( values=pal ) +
    ggrepel::geom_text_repel( aes(label = Abbrev), show.legend=FALSE ) +
    geom_abline( slope = 1, linetype = "dashed", color = "gray40" ) +
    xlab( "Replicate 1" ) + ylab( "Replicate 2" )
```

## Evaluation of performance

```{r}
R <- P %>% select( ABCpred, Drug, pubchem_id ) %>% inner_join(S)
```
