---
title: "Overview of raw data"
output: html_document
---

## Accessing the data

The package contains the training and test datasets described in the
paper. You can access them directly via the `data()` function after
loading the package:

```{r}
library( ABCmonster )
data( MACCSbinary )
## The data is now available as MACCSbinary
```

Looking at the first few rows and columns gives the general idea of how
the data is structured.

```{r}
dim( MACCSbinary )
head( MACCSbinary[,1:6] )
```

The first column, `Label`, contains annotations for each drug
specifying whether the drug is more (`"Sensitive"`) or less (`"Resistant"`)
efficacious in the ABC-16 strain, relative to the parental strain.
Test data is located in the last 24 rows of `MACCSbinary` and contains `NA`s in the
`Label` column:

```{r}
tail( MACCSbinary[,1:6] )
```

Drug identity is captured in the second (`Drug`) and third
(`pubchem_id`) columns. All remaining columns contain molecular
features, which in this case consist of binary calls on whether a
particular MACCS key is present in a drug's structure.

## Visualizing the data

Let's begin by plotting a heatmap that presents and overview of the
entire dataset. We will make extensive use of
[tidy](https://www.jstatsoft.org/article/view/v059i10) data
manipulation. As the first step, we load the appropriate libraries and
define a palette for subsequent plots.

```{r, message=FALSE}
library( tidyverse )
library( plotly )
ABCpal <- c("Resistant" = "tomato", "Sensitive" = "steelblue")
```

Isolate the training data by selecting the set of rows with non-`NA`
labels. `HH` will contain the matrix to be plotted as the heatmap,
while `Annot` will store labels that will be displayed as an additional
annotation bar at the top of the heatmap. The plotting is done using
the [pheatmap package](https://github.com/raivokolde/pheatmap).

```{r fig.align="center"}
Xtrain <- MACCSbinary %>% filter( !is.na(Label) )
HH <- Xtrain %>% select(-Label, -Drug) %>% column_to_rownames("pubchem_id") %>% as.matrix
Annots <- select( Xtrain, pubchem_id, Label ) %>% column_to_rownames( "pubchem_id" )
pheatmap::pheatmap( t(HH), show_rownames=FALSE, show_colnames=FALSE,
         annotation_col=Annots, annotation_names_col=FALSE, legend=FALSE,
         color=c("gray90","gray30"), annotation_colors=list(Label=ABCpal) )
```

The version of this figure in the paper includes additional
post-processing consisting of leaf order optimization and grouping of rows.

### Dimensionality reduction

Lastly, we can perform dimensionality reduction to explore the general
trends in the data.

```{r eval=FALSE}
## Compute PCA
P <- Xtrain %>% select( -Label, -Drug, -pubchem_id ) %>% prcomp() %>%
    broom::augment( Xtrain ) %>% select( Label, PC1=.fittedPC1, PC2=.fittedPC2 )

## Plot the projection onto the first two principal components
gg <- ggplot( P, aes(x=PC1, y=PC2, color=Label) ) + geom_point() +
    theme_bw() + scale_color_manual( values=ABCpal )
ggplotly(gg)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
P <- Xtrain %>% select( -Label, -Drug, -pubchem_id ) %>% prcomp() %>%
    broom::augment( Xtrain ) %>% rename( PC1 = .fittedPC1, PC2 = .fittedPC2 )
gg <- ggplot( P, aes(x=PC1, y=PC2, color=Label, lbl1=Drug, lbl2=pubchem_id) ) +
    geom_point() + theme_bw() + scale_color_manual( values=ABCpal, guide=FALSE )
ggplotly(gg, tooltip=c("lbl1", "lbl2"), width=600) %>%
    add_annotations( text="Label", xref="paper", yref="paper",
                    x=1.02, xanchor="left",
                    y=0.8, yanchor="bottom",    # Same y as legend below
                    legendtitle=TRUE, showarrow=FALSE ) %>%
    layout( legend=list(y=0.8, yanchor="top" ) ) %>%
    htmltools::div( align="center" )
```

We observe that the unsupervised analysis is unable to distinguish
between the class labels. In the next vignette, we will investigate this data in the
context of [supverised methods](sv.html).

## Additional Feature Spaces

The package includes three additional feature representations of the same
400 drugs: 1) `MACCScount`, which captures the number of occurrences
of each MACCS fingerprint (as opposed to just their presence/absence,
as in `MACCSbinary`); 2) `Morgan`, which captures ECFP/Morgan
representation; and 3) `PChem`, which consists of physico-chemical
properties. All three matrices are in the same format as
`MACCSbinary`, including having the `Label` column designate class
assignment (`Sensitive` or `Resistant`) for training data and `NA` for
test samples.

We can load and examine the data using the same functions as those
outlined in the previous sections. Here's a couple of additional examples:
```{r}
## Load all three datasets
data( MACCScount, Morgan, PChem )

## Identify the highest number of occurrences of MACCS keys
## List the top keys along with which drugs they occur in
MACCScount %>% gather( Key, nOccur, -Label, -Drug, -pubchem_id ) %>% group_by( Key ) %>%
    filter( nOccur == max(nOccur) ) %>% arrange( desc(nOccur) ) %>% ungroup %>% head

## Count the total number of occurrences of Morgan keys
## Display the most commonly-occurring keys
Morgan %>% gather( Key, nOccur, -Label, -Drug, -pubchem_id ) %>% group_by( Key ) %>%
    summarize( nTotal = sum(nOccur) ) %>% arrange( desc(nTotal) ) %>% head

## We may be interested in normalizing each physicochemical property
##   to have zero mean and standard deviation of one
PChemNorm <- PChem %>% mutate_at( vars(-Label, -Drug, -pubchem_id), ~(.x - mean(.x))/sd(.x) )
summary( PChemNorm$SlogP )
sd( PChemNorm$SlogP )
```
